---
title: "Untitled"
execute: 
  freeze: true
format:
  html:
    highlight: true
    code-fold: show
    code-tools: true
    highlight-style: github
---

# Nifty Features with Quarto

## Callouts

:::{.callout-note}
Note that there are five types of callouts, including:
`note`, `warning`, `important`, `tip`, and `caution`.
:::

:::{.callout-tip}
## Tip With Caption

This is an example of a callout with a caption.
:::

:::{.callout-caution collapse="true"}
## Expand To Learn About Collapse

This is an example of a 'folded' caution callout that can be expanded by the user. You can use `collapse="true"` to collapse it by default or `collapse="false"` to make a collapsible callout that is expanded by default.
:::

### Appearance

::: {.callout-note appearance="simple"}

## Pay Attention

Using callouts is an effective way to highlight content that your reader give special consideration or attention. You can set this globally `callout-appearance: simple`.

:::

### Icon

::: {.callout-note icon=false}

## Pay Attention

Using callouts is an effective way to highlight content that your reader give special consideration or attention `callout-icon: false;`.

:::

## Margin Figures

```{r}
#| label: fig-mtcars
#| fig-cap: "MPG vs horsepower, colored by transmission."
#| column: margin

library(ggplot2)
mtcars2 <- mtcars
mtcars2$am <- factor(
  mtcars$am, labels = c('automatic', 'manual')
)
ggplot(mtcars2, aes(hp, mpg, color = am)) +
  geom_point() +  geom_smooth(formula = y ~ x, method = "loess") +
  theme(legend.position = 'bottom')
```

### Margin Tables

```{r}
#| column: margin

knitr::kable(
  mtcars[1:3, 1:3]
)
```

### Other Content

::: {.column-margin}
We know from *the first fundamental theorem of calculus* that for $x$ in $[a, b]$:


$$\frac{d}{dx}\left( \int_{a}^{x} f(u)\,du\right)=f(x).$$

:::